Global land characterisation using cover fractions at 100m resolution

### 2020-08-05 article/article.tex
Final changes

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
%\newacronym{CGLS}{CGLS}{Copernicus Global Land Services}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{WUR}{WUR}{Wageningen University \& Research}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NDMI}{NDMI}{Normalised Difference Moisture Index}
\newacronym{NIRv}{NIR\textsubscript{v}}{\gls{NIR} of vegetation}
\newacronym{EVI}{EVI}{Enhanced Vegetation Index}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SIF}{SIF}{sun-induced chlorophyll fluorescence}
\newacronym{GPP}{GPP}{gross primary productivity}
\newacronym{MSI}{MSI}{multi-spectral instrument}
\newacronym{SAR}{SAR}{synthetic aperture radar}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{classification and regression tree}
\newacronym{GLM}{GLM}{general linear regression model}
\newacronym{LCCS}{LCCS}{UN Land Cover Classification Scheme}
\newacronym{GDAL}{GDAL}{Geospatial Data Abstraction Library}
%\makenoidxglossaries
% Disable abbreviating RF for now [text={random forest}]
%\glsunset{RF}

% Use \begin{equation} for maths

@@

\begin{keyword}
Land cover fraction mapping\sep machine learning\sep random forest\sep Cubist\sep neural networks\sep fuzzy nearest centroid\sep lasso regression\sep PROBA-V
\end{keyword}

@@

For climate data, WorldClim 2.0 30 second product \citep{worldclim2} was used.
It includes monthly temperature, precipitation, solar radiation, wind speed and water vapour pressure data.
In addition, 19 bioclimatic parameters were calculated from the data, using the dismo package \citep{dismo}.
Some additional biophysical parameters were also calculated, namely all of the climate data during the coldest, warmest, driest and wettest months of the year at each location, as well as yearly averages of each of the original parameters.

@@

In total, 313 covariates were generated in pre-processing.
However, many of them were collinear with one another.
Multicollinearity prolongs training time for machine learning models and leads to unreliable coefficient estimation and increased error variance for linear models.
Thus, variable selection was used to remove collinear covariates before predicting the land cover fractions.
Covariates that had a Pearson's correlation coefficient $r$ \citep{pearson_notes_1895} above 0.9 were excluded in an iterative process.
After that, covariates with a Spearman's rank correlation $\rho$ \citep{spearman1904rank} above 0.9 were likewise excluded.

@@

\Gls{FNC} regression \citep{keller_fuzzy_1985}, also called fuzzy nearest prototype, fuzzy $c$-means or fuzzy $k$-means, is a simple regression method where the pixel's membership of a class is determined by the distance of the pixel from each class centroid in feature space.
This method always produces output that sums to 100\%.
In this study, we used an implementation adapted from the package GSIF \citep{hengl2004fuzzycmeans}.
We tested both using a logistic regression to estimate the class centroids, as well as using a weighted average of the training data to do so, and ended up using the latter method as the result was more accurate.

@@

In this study, we compared three groups of algorithms for land cover fraction mapping and created a demonstation map showing the global distribution of land cover in fraction layers, using the best performing model according to \gls{RMSE}.
We also introduced a multi-step approach to tackle the issue of accurately predicting the extreme fraction values of 0\% and 100\%.
The best performing models were based on \gls{CART} tree ensembles, with \gls{RF} regression performing the best overall.
A single-step \gls{RF} regression model using majority voting resulted in the best results for RMSE (17.3\%) and \gls{NSE} (0.66).
Using median voting combined with the three-step approach obtained the best results for \gls{MAE} (7.9\%) and \gls{SCM} \gls{OA} (72±2\%).
Remote sensing covariates were the most important in the models, although all other types of covariates (climate, soil, terrain) also contributed to the model significantly.

@@

% Converted from role-centric to person-centric as per Elsevier guidelines: https://www.elsevier.com/authors/journal-authors/policies-and-ethics/credit-author-statement
\minisection{Author Contributions}
\textbf{Dainius Masiliūnas}: Methodology, Formal Analysis, Software, Writing – Original Draft, Data Curation. \textbf{Nandin-Erdene Tsendbazar}: Investigation, Data Curation, Supervision, Conceptualisation, Writing – review \& editing. \textbf{Martin Herold}: Conceptualisation, Project Administration, Supervision, Methodology. \textbf{Myroslava Lesiv}: Investigation, Data Curation. \textbf{Marcel Buchhorn}: Writing – review \& editing. \textbf{Jan Verbesselt}: Conceptualisation, Methodology, Supervision, Writing – review \& editing.

\minisection{Acknowledgements} We would like to thank VITO and Terrascope for providing us with computational facilities to access and process {PROBA-V} data, and IIASA for providing the model training dataset. In addition, we thank Bruno Smets for ideas and support from the \gls{CGLS-LC100} project, Benjamin Kellenberger for help on tuning \gls{NN} parameterisation, as well as Linlin Li and Tomislav Hengl for suggestions on additional machine learning methods to test.

%\minisection{Conflicts of Interest} The authors declare no conflict of interest.

@@

>>>>

% Acronyms - cite with \ac{}
\newacronym{GNSS}{GNSS}{Global Navigation Satellite System}
\newacronym{OA}{OA}{overall accuracy}
\newacronym{UA}{UA}{user accuracy}
\newacronym{PA}{PA}{producer accuracy}
\newacronym{RMSE}{RMSE}{root mean squared error}
\newacronym{MAE}{MAE}{mean absolute error}
\newacronym{ME}{ME}{mean error}
\newacronym{NSE}{NSE}{Nash–Sutcliffe model efficiency coefficient}
\newacronym{SDG}{SDG}{Sustainable Development Goal}
\newacronym{LC-CCI}{LC-CCI}{ESA Climate Change Initiative Land Cover}
\newacronym{CGLS-LC100}{CGLS-LC100}{Copernicus Global Land Services 100 m Land Cover}
\newacronym{FROM-GLC}{FROM-GLC10}{Finer Resolution Observation and Monitoring Global Land Cover}
\newacronym{IIASA}{IIASA}{International Institute for Applied Systems Analysis}
\newacronym{WUR}{WUR}{Wageningen University \& Research}
\newacronym{SMA}{SMA}{spectral mixture analysis}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{MLP}{MLP}{multi-layer perceptron}
\newacronym{NN}{NN}{neural network}
\newacronym{LOESS}{LOESS}{locally estimated scatterplot smoothing}
\newacronym[firstplural={vegetation indices (VIs)}]{VI}{VI}{vegetation index}
\newacronym{GLSDEM}{GLSDEM}{Global Land Survey Digital Elevation Model}
\newacronym{TPI}{TPI}{Terrain Position Index}
\newacronym{OSAVI}{OSAVI}{Optimised Soil-Adjusted Vegetation Index}
\newacronym{NDVI}{NDVI}{Normalised Difference Vegetation Index}
\newacronym{NDMI}{NDMI}{Normalised Difference Moisture Index}
\newacronym{NIRv}{NIR\textsubscript{v}}{\gls{NIR} of vegetation}
\newacronym{EVI}{EVI}{Enhanced Vegetation Index}
\newacronym{NIR}{NIR}{near infra-red}
\newacronym{SIF}{SIF}{sun-induced chlorophyll fluorescence}
\newacronym{GPP}{GPP}{gross primary productivity}
\newacronym{MSI}{MSI}{multi-spectral instrument}
\newacronym{SAR}{SAR}{synthetic aperture radar}
\newacronym{SCM}{SCM}{subpixel confusion-uncertainty matrix}
\newacronym{FNC}{FNC}{fuzzy nearest centroid}
\newacronym{PLS}{PLS}{partial least squares}
\newacronym{MLR}{MLR}{multinomial logistic regression}
\newacronym{RF}{RF}{random forest}
\newacronym{TOC}{TOC}{Top-of-Canopy}
\newacronym{CART}{CART}{classification and regression tree}
\newacronym{GLM}{GLM}{general linear regression model}
\newacronym{LCCS}{LCCS}{UN Land Cover Classification Scheme}
\newacronym{GDAL}{GDAL}{Geospatial Data Abstraction Library}
%\makenoidxglossaries

@@

\begin{keyword}
PROBA-V\sep global land cover mapping\sep land cover fraction mapping\sep time series analysis\sep machine learning\sep random forest\sep Cubist\sep support vector regression\sep neural network\sep spatial accuracy
\end{keyword}

@@

For climate data, WorldClim 2.0 30 second product \citep{worldclim2} was used.
It includes monthly temperature, precipitation, solar radiation, wind speed and water vapour pressure data.
In addition, 19 bioclimatic parameters were calculated from the data, using the \cran{dismo} package \citep{dismo}.
Some additional biophysical parameters were also calculated, namely all of the climate data during the coldest, warmest, driest and wettest months of the year at each location, as well as yearly averages of each of the original parameters.

@@

In total, 313 covariates were generated in preprocessing.
However, many of them were collinear with one another.
Multicollinearity prolongs training time for machine learning models and leads to unreliable coefficient estimation and increased error variance for linear models.
Thus, variable selection was used to remove collinear covariates before predicting the land cover fractions.
Covariates that had a Pearson's correlation coefficient $r$ \citep{pearson_notes_1895} above 0.9 were excluded in an iterative process.
After that, covariates with a Spearman's rank correlation $\rho$ \citep{spearman1904rank} above 0.9 were likewise excluded.

@@

\Gls{FNC} regression \citep{keller_fuzzy_1985}, also called fuzzy nearest prototype, fuzzy $c$-means or fuzzy $k$-means, is a simple regression method where the pixel's membership of a class is determined by the distance of the pixel from each class centroid in feature space.
This method always produces output that sums to 100\%.
In this study, we used an implementation adapted from the package \cran{GSIF} \citep{hengl2004fuzzycmeans}.
We tested both using a logistic regression to estimate the class centroids, as well as using a weighted average of the training data to do so, and ended up using the latter method as the result was more accurate.

@@

In this study, we compared three groups of algorithms for land cover fraction mapping and created a demonstration map showing the global distribution of land cover in fraction layers, using the best performing model according to \gls{RMSE}.
We also introduced a multi-step approach to tackle the issue of accurately predicting the extreme fraction values of 0\% and 100\%.
The best performing models were based on \gls{CART} tree ensembles, with \gls{RF} regression performing the best overall.
A single-step \gls{RF} regression model using majority voting resulted in the best results for RMSE (17.3\%) and \gls{NSE} (0.66).
Using median voting combined with the three-step approach obtained the best results for \gls{MAE} (7.9\%) and \gls{SCM} \gls{OA} (72±2\%).
Remote sensing covariates were the most important in the models, although all other types of covariates (climate, soil, terrain) also contributed to the model significantly.

@@

% Converted from role-centric to person-centric as per Elsevier guidelines: https://www.elsevier.com/authors/journal-authors/policies-and-ethics/credit-author-statement
%\minisection{Author Contributions}
%\textbf{Dainius Masiliūnas}: Methodology, Formal Analysis, Software, Writing – Original Draft, Data Curation. \textbf{Nandin-Erdene Tsendbazar}: Investigation, Data Curation, Supervision, Conceptualisation, Writing – review \& editing. \textbf{Martin Herold}: Conceptualisation, Project Administration, Supervision, Methodology. \textbf{Myroslava Lesiv}: Investigation, Data Curation. \textbf{Marcel Buchhorn}: Writing – review \& editing. \textbf{Jan Verbesselt}: Conceptualisation, Methodology, Supervision, Writing – review \& editing.

\minisection{Acknowledgements} We would like to thank VITO and Terrascope for providing us with computational facilities to access and process {PROBA-V} data, and IIASA for providing the model training dataset. In addition, we thank Bruno Smets for ideas and support from the \gls{CGLS-LC100} project, Benjamin Kellenberger for help on tuning \gls{NN} parametrisation, as well as Linlin Li and Tomislav Hengl for suggestions on additional machine learning methods to test.

@@

%\listoffigures
